### 2019-04-30 papers/haskell19/6-casestudies.tex
Case studies

As recently analyzed by \citeauthor{KiriyamaOptimizingDTC}, this slowdown is
expected to be linear in the depth of our representation type
\cite{KiriyamaOptimizingDTC}.
%
Hence, one can drastically reduce the runtime overhead by associating each
|oplus| operator in a balanced fashion.
%
So, for instance, instead of writing |(f oplus'' g oplus'' h oplus'' i)|, which
is implicitly parsed as |(f oplus'' (g oplus'' (h oplus'' i)))|; we can
associate constructions as |((f oplus'' g) oplus'' (h oplus'' i))|, thus
reducing the depth of our representation from four to three levels and, in
general, from a $\mathcal{O}(n)$ to a $\mathcal{O}(log(n))$ complexity in the
runtime overhead, where $n$ is the amount of constructions under consideration.
%
To apply this optimization automatically, and relying on more sophisticated
techniques like prefix-free encodings to obtain optimal representations, is a
undergoing path of our future work.

@@

\begin{table}[H]
  \begin{tabular}{||c||c||c||c||c||}
    \hline
    Case Study & |Con|s & |Pat|s & |Fun|s & Total Constructions \\ \hline
    \hline
    RBT        & 2      & 6      & 5      & 13                  \\ \hline
    SExp       & 6      & 9      & -      & 15                  \\ \hline
    Html       & 4      & -      & 132    & 136                 \\ \hline
  \end{tabular}
\end{table}

For each case study, we benchmarked the execution time required to generate 1000
random values using both manually-written concrete generators, and modular
generators obtained using our approach.
%
Additionally, our modular generators were tested using both linear and balanced
generation specifications.
%
Figure \ref{fig:times} illustrates the relative execution time of each case
study, with respect to each manually-written random generator---we encourage the
reader to obtain a colored version of this work.

\begin{figure}[b]
  \centering
  \begin{tikzpicture}
    \begin{axis}
      [ height=6cm
      , width=9cm
      , ybar
      , ymajorgrids=true
      , ytick={0,1,...,5}
      , enlargelimits=0.25
      , enlarge x limits=0.25
      , /pgf/bar width=17pt
      , legend style={
        at={(0.5,-0.15)},
        anchor=north,legend columns=3}
      , legend image code/.code={%
        \draw[#1, draw=none] (0cm,-0.1cm) rectangle (0.3cm,0.2cm); }
      , ylabel={Relative Generation Time}
      , symbolic x coords={RBT, SExp, Html}
      , xtick=data
      % , nodes near coords
      , nodes near coords align={vertical}
      ]
      \addplot[fill, red!30]   coordinates {(SExp,1)    (RBT,1)    (Html,1)};
      \addplot[fill, green!30] coordinates {(SExp,2.31) (RBT,3)    (Html,4.88)};
      \addplot[fill, blue!30]  coordinates {(SExp,1.62) (RBT,2.28) (Html,1.70)};
      \legend{Manual, Modular/Linear, Modular/Balanced}
    \end{axis}
  \end{tikzpicture}
  \caption{Generation time comparison between concrete and modular generators.}
  \label{fig:times}
\end{figure}

>>>>

As recently analyzed by \citeauthor{KiriyamaOptimizingDTC}, this slowdown is
expected to be linear in the depth of our representation type
\cite{KiriyamaOptimizingDTC}.
%
Hence, one can drastically reduce the runtime overhead by associating each
|oplus| operator in a balanced fashion.
%
So, for instance, instead of writing |(f oplus'' g oplus'' h oplus'' i)|, which
is implicitly parsed as |(f oplus'' (g oplus'' (h oplus'' i)))|; we can
associate constructions as |((f oplus'' g) oplus'' (h oplus'' i))|, thus
reducing the depth of our representation from four to three levels and, in
general, from a $\mathcal{O}(n)$ to a $\mathcal{O}(log(n))$ complexity in the
runtime overhead, where $n$ is the amount of constructions under consideration.

@@

\begin{table}[H]
  \begin{tabular}{||c||c||c||c||c||}
    \hline
    Case Study & |Con|s & |Pat|s & |Fun|s & Total Constructions \\ \hline
    \hline
    RBT        & 2      & 6      & 5      & 13                  \\ \hline
    SExp       & 6      & 9      & -      & 15                  \\ \hline
    HTML       & 4      & -      & 132    & 136                 \\ \hline
  \end{tabular}
\end{table}

We benchmarked the execution time required to generate and fully evaluate 1000
random values corresponding to each case study, comparing both manually-written
concrete generators and modular generators obtained using our approach.
%
For this purpose, we used the \emph{Criterion} benchmarking tool for Haskell,
and limited the maximum depth of the generated values to five levels.
%
Additionally, our modular generators were tested using both linear and balanced
generation specifications.
%
Figure \ref{fig:times} illustrates the relative execution time of each case
study, with respect to each manually-written random generator---we encourage the
reader to obtain a colored version of this work.

\begin{figure}[b]
  \centering
  \begin{tikzpicture}
    \begin{axis}
      [ height=6cm
      , width=9cm
      , ybar
      , ymajorgrids=true
      , ytick={0,1,...,5}
      , ymax=4.5
      , enlargelimits=0.25
      , enlarge x limits=0.25
      , /pgf/bar width=17pt
      , legend style={
        at={(0.5,-0.15)},
        anchor=north,legend columns=3}
      , legend image code/.code={%
        \draw[#1, draw=none] (0cm,-0.1cm) rectangle (0.3cm,0.2cm); }
      , ylabel={Relative Generation Time}
      , symbolic x coords={RBT, SExp, HTML}
      , xtick=data
      % , nodes near coords
      , nodes near coords align={vertical}
      ]
      \addplot[fill, red!30]   coordinates {(SExp,1)    (RBT,1)    (HTML,1)};
      \addplot[fill, green!30] coordinates {(SExp,2.31) (RBT,3)    (HTML,4.88)};
      \addplot[fill, blue!30]  coordinates {(SExp,1.62) (RBT,2.28) (HTML,1.70)};
      \legend{Manual, Modular/Linear, Modular/Balanced}
    \end{axis}
  \end{tikzpicture}
  \caption{Generation time comparison between concrete and modular generators.}
  \label{fig:times}
\end{figure}

As it can be observed, our approach suffers from a noticeable runtime overhead,
specially when considering the HTML case study with a large amount of
construction involved at the generation process.
%
However, by balancing our representation types we found that the generation
performance gets dramatically improved.
%
To apply this optimization automatically, and relying on more sophisticated
techniques like prefix-free encodings to obtain optimal representations, is a
undergoing path of our future work.

On the other hand, it has been argued that the generation time is often not
substantial with respect to the rest of the testing process, especially when
testing complex properties over monadic code, as well as using randomly
generated values for penetration testing.

Nonetheless, we consider that these results are in turn very encouraging, given
that the flexibility obtained from using our approach does not produce severe
slowdowns in practice, nor explodes more than linearly as we increase the
complexity of our random generators.
